Parallel Tractability of Ontology\\ Materialization: Technique and Practice

### 2018-01-31 parallel-tractability-J-v9/appendix.tex
Finished the proof-reading of the optimized algorithm section

\emph{Inductive case}.
According to the induction hypothesis, the partial graph constructed after
the $i^{th}$ ($i>1$) iteration is a materialization graph, denoted by $\mathcal{G}^{i}$.
We have to show that the partial graph constructed after the $i+1^{th}$ iteration is
also a materialization graph, denoted by $\mathcal{G}^{i+1}$.
The partial graph $\mathcal{G}^{i+1}$ is updated in $\mathsf{A}_{opt}$ by performing the step $(\textbf{\romannumeral3})$
of Algorithm~$\mathsf{Opt}$.
Suppose that \texttt{rch}$(B_k,H)\in S_{\tiny rch}$ is checked.
It corresponds to the rule instantiation $B_1,..,B_k,..,B_n\rightarrow H$.
Algorithm~$\mathsf{Opt}$ next checks that there exists a node $B'$ such that
\texttt{rch}$(B',B_k)\in S^*_{\tiny rch}$
and $B'$ is in $\mathcal{G}$. This means that $B_k$ and $H$ are derivable.
The algorithm then adds new nodes $B_k$ and $H$ to $\mathcal{G}^{i}$ in three cases.
(1) if $H$ is not in $\mathcal{G}^{i}$ while $B_k$ is in $\mathcal{G}^{i}$,
then $H$ is added to $\mathcal{G}^{i+1}$, and the edges $e(B_1, H),...,e(B_n, H)$ are created.
(2) when neither of $H$ and $B_k$ is in $\mathcal{G}^{i}$,
then $H$ and $B_k$ are added to $\mathcal{G}^{i+1}$, and the edges $e(B_1, H),...,e(B_n, H)$ are also created.
In the above two cases, $H$ is a new node. Thus,
$\mathcal{G}^{i+1}$ is acyclic.
(3) if $H$ is in $\mathcal{G}^{i}$ and $H$ has no parent,
$B_k$ is added to $\mathcal{G}^{i+1}$, and the edges $e(B_1, H),...,e(B_n, H)$ are also created.
For the case that $H$ is in $\mathcal{G}^{i}$ and $H$ has parents,
the algorithm does nothing.
Thus, $\mathcal{G}^{i+1}$ is acyclic and satisfies the definition of materialization graph.

@@

\noindent\emph{Proof}: similar to the proof of Lemma~\ref{lemma:a3},
this theorem is proved in two stages:
(1) the graph $\mathcal{G}$ returned by $\mathsf{A}_{prc}$ is a materialization graph;
(2) $\mathcal{G}$ is a complete materialization graph.
One can show that (1) holds by performing the same induction on the iterations of \ref{alg3:updateG} of $\mathsf{A}_{opt}$
(see the proof of Lemma~\ref{lemma:a3}). This is because that
such an induction is performed regardless of how the relation $S_{\textit{\tiny rch}}$ is computed.

To prove that (2) holds, we conduct the induction used in the proof for Lemma~\ref{lemma:a1}.
We aim to show that all the ground atoms in $T_R^{i+1}(\textbf{I})$ have to be added to $\mathcal{G}$,
with the induction hypothesis that the ground atoms in $T_R^{i}(\textbf{I})$ are in $\mathcal{G}$.
Suppose the ground atoms in $T_R^{i}(\textbf{I})$ have been added to $\mathcal{G}$ by
performing $\mathsf{A}_{prc}$.
For each atom $\alpha$ in $T_R^{i+1}(\textbf{I})$, there has to be a relation $(\beta,\alpha)$
in $S_{\textit{\tiny rch}}$ obtained in some iteration; further
$\beta$ is already in $\mathcal{G}$ according to the inductive hypothesis. Thus,
$\alpha$ has to be added to $\mathcal{G}$ by applying \ref{alg4:updateG}
of $\mathsf{A}_{prc}$.\hfill$\Box$

@@

We first consider Stage~SCM.
Stage~SCM (the simple concept materialization) is conducted on axioms of two forms $A\sqsubseteq B$
and $\exists R.A\sqsubseteq B$, which correspond to datalog rules of the forms $A(x)\rightarrow B(x)$
and $R(x,y),A(y)\rightarrow B(x)$ respectively. Since role assertions are supposed to be fixed
during in Stage~SCM, role $R$ in each rule of the form $R(x,y),A(y)\rightarrow B(x)$
can be viewed as an EDB predicate. This further means that,
for each atom of the form $A(a)$ where $A$ is a simple concept,
$A(a)$ has an SWD path in the first iteration of $\mathsf{A}_{prc}$.
Recall Algorithm~$\mathsf{Prc}$. We use an induction to show that
such an SWD path of $A(a)$ can be determined by $S_{\textit{\tiny rch}}$.
The atom $A(a)$ can be derived through either of the following rule instantiations:

\begin{enumerate}[leftmargin=12ex]
\item[Case~1] $B(a)\rightarrow A(a)$, where $\mathcal{O}\models B\sqsubseteq_* A$. We have
    that \texttt{rch}$(B(a),A(a))$ is in $S_{\textit{\tiny rch}}$ according to Algorithm~$\mathsf{Prc}$.
    By the induction hypothesis, the existence of the SWD path for $B(a)$ can be determined.
    Thus, the existence of the SWD path of $A(a)$ can also be determined by $S_{\textit{\tiny rch}}$.

\item[Case~2] $R(a,b),B(b)\rightarrow A(a)$ where $\exists R.B\sqsubseteq A\in\mathcal{T}$.
    According to Algorithm~$\mathsf{Prc}$ and the fact that $R(a,b)$ has to be derived,
    we have that \texttt{rch}$(B(b),$ $A(a))$ is in $S_{\textit{\tiny rch}}$.
    By the induction hypothesis, the existence of the SWD path for $B(b)$ can be determined.
    Thus, the existence of the SWD path of $A(a)$ can also be determined.
\end{enumerate}

@@

\begin{enumerate}[leftmargin=12ex]
\item[Case~3] $A_1(a),A_2(a)\rightarrow A(a)$ where $A_1\sqcap A_2\sqsubseteq A\in\mathcal{T}$.
    If $A_1(a)$ is derived where $A_1$ is a simple concept,
    we have that \texttt{rch}$(A_2(a),A(a))$ is in $S_{\textit{\tiny rch}}$
    according to Algorithm~$\mathsf{Prc}$.
    By the induction hypothesis, the existence of the SWD path for $A_2(a)$ can be determined.
    Thus, the existence of the SWD path of $A(a)$ can also be determined.
    It is similar for the case where $A_2(a)$ is derived and $A_2$ is a simple concept.
\end{enumerate}

@@

\begin{enumerate}[leftmargin=12ex]
\item[Case~4] $S(a,b)\rightarrow R(a,b)$ where $\mathcal{O}\models S\sqsubseteq_* R$.
    We have that \texttt{rch}$(S(a,b),R(a,b))$ is in $S_{\textit{\tiny rch}}$
    according to Algorithm~$\mathsf{Prc}$.
    By the induction hypothesis, the existence of the SWD path for $S(a,b)$ can be determined.
    Thus, the existence of the SWD path of $R(a,b)$ can also be determined.

\item[Case~5] $S(b,a)\rightarrow R(a,b)$ where $\mathcal{O}\models S\sqsubseteq_* R^-$.
    We have that \texttt{rch}$(S(b,a),R(a,b))$ is in $S_{\textit{\tiny rch}}$
    according to Algorithm~$\mathsf{Prc}$.
    By the induction hypothesis, the existence of the SWD path for $S(b,a)$ can be determined.
    Thus, the existence of the SWD path of $R(a,b)$ can also be determined.

\item[Case~6] $R_1(a,c),R_2(c,b)\rightarrow R(a,b)$ where $R_1\circ R_2\sqsubseteq R\mathcal{R}$.
    If $R_1(a,c)$ is derived where $R_1$ is a simple role,
    we have that \texttt{rch}$(R_2(c,b),R(a,b))$ is in $S_{\textit{\tiny rch}}$
    according to Algorithm~$\mathsf{Prc}$.
    By the induction hypothesis, the existence of the SWD path for $R_2(c,b)$ can be determined.
    Thus, the existence of the SWD path of $R(a,b)$ can also be determined.
    It is similar for the case where $R_2(c,b)$ is derived and $R_2$ is a simple role.
\end{enumerate}

>>>>

\emph{Inductive case}.
According to the induction hypothesis, the partial graph constructed after
the $i^{th}$ ($i>1$) iteration is a materialization graph, denoted by $\mathcal{G}^{i}$.
We have to show that the partial graph constructed after the $i+1^{th}$ iteration is
also a materialization graph, denoted by $\mathcal{G}^{i+1}$.
The partial graph $\mathcal{G}^{i+1}$ is updated in $\mathsf{A}_{opt}$ by performing the step $(\textbf{\romannumeral3})$
of Algorithm~$\mathsf{Opt}$.
Suppose that \texttt{rch}$(B_k,H)\in S_{\textit{\!\tiny rch}}$ is checked.
It corresponds to the rule instantiation $B_1,..,B_k,..,B_n\rightarrow H$.
Algorithm~$\mathsf{Opt}$ next checks that there exists a node $B'$ such that
\texttt{rch}$(B',B_k)\in S^*_{\!\tiny rch}$
and $B'$ is in $\mathcal{G}$. This means that $B_k$ and $H$ are derivable.
The algorithm then adds new nodes $B_k$ and $H$ to $\mathcal{G}^{i}$ in three cases.
(1) if $H$ is not in $\mathcal{G}^{i}$ while $B_k$ is in $\mathcal{G}^{i}$,
then $H$ is added to $\mathcal{G}^{i+1}$, and the edges $e(B_1, H),...,e(B_n, H)$ are created.
(2) when neither of $H$ and $B_k$ is in $\mathcal{G}^{i}$,
then $H$ and $B_k$ are added to $\mathcal{G}^{i+1}$, and the edges $e(B_1, H),...,e(B_n, H)$ are also created.
In the above two cases, $H$ is a new node. Thus,
$\mathcal{G}^{i+1}$ is acyclic.
(3) if $H$ is in $\mathcal{G}^{i}$ and $H$ has no parent,
$B_k$ is added to $\mathcal{G}^{i+1}$, and the edges $e(B_1, H),...,e(B_n, H)$ are also created.
For the case that $H$ is in $\mathcal{G}^{i}$ and $H$ has parents,
the algorithm does nothing.
Thus, $\mathcal{G}^{i+1}$ is acyclic and satisfies the definition of materialization graph.

@@

\noindent\emph{Proof}: similar to the proof of Lemma~\ref{lemma:a3},
this theorem is proved in two stages:
(1) the graph $\mathcal{G}$ returned by $\mathsf{A}_{prc}$ is a materialization graph;
(2) $\mathcal{G}$ is a complete materialization graph.
One can show that (1) holds by performing the same induction on the iterations of \ref{alg3:updateG} of $\mathsf{A}_{opt}$
(see the proof of Lemma~\ref{lemma:a3}). This is because that
such an induction is performed regardless of how the relation $S_{\textit{\!\tiny rch}}$ is computed.

To prove that (2) holds, we conduct the induction used in the proof for Lemma~\ref{lemma:a1}.
We aim to show that all the ground atoms in $T_R^{i+1}(\textbf{I})$ have to be added to $\mathcal{G}$,
with the induction hypothesis that the ground atoms in $T_R^{i}(\textbf{I})$ are in $\mathcal{G}$.
Suppose the ground atoms in $T_R^{i}(\textbf{I})$ have been added to $\mathcal{G}$ by
performing $\mathsf{A}_{prc}$.
For each atom $\alpha$ in $T_R^{i+1}(\textbf{I})$, there has to be a relation $(\beta,\alpha)$
in $S_{\textit{\!\tiny rch}}$ obtained in some iteration; further
$\beta$ is already in $\mathcal{G}$ according to the inductive hypothesis. Thus,
$\alpha$ has to be added to $\mathcal{G}$ by applying \ref{alg4:updateG}
of $\mathsf{A}_{prc}$.\hfill$\Box$

@@

We first consider Stage~SCM.
Stage~SCM (the simple concept materialization) is conducted on axioms of two forms $A\sqsubseteq B$
and $\exists R.A\sqsubseteq B$, which correspond to datalog rules of the forms $A(x)\rightarrow B(x)$
and $R(x,y),A(y)\rightarrow B(x)$ respectively. Since role assertions are supposed to be fixed
during in Stage~SCM, role $R$ in each rule of the form $R(x,y),A(y)\rightarrow B(x)$
can be viewed as an EDB predicate. This further means that,
for each atom of the form $A(a)$ where $A$ is a simple concept,
$A(a)$ has an SWD path in the first iteration of $\mathsf{A}_{prc}$.
Recall Algorithm~$\mathsf{Prc}$. We use an induction to show that
such an SWD path of $A(a)$ can be determined by $S_{\textit{\!\tiny rch}}$.
The atom $A(a)$ can be derived through either of the following rule instantiations:

\begin{enumerate}[leftmargin=12ex]
\item[Case~1] $B(a)\rightarrow A(a)$, where $\mathcal{O}\models B\sqsubseteq_* A$. We have
    that \texttt{rch}$(B(a),A(a))$ is in $S_{\textit{\!\tiny rch}}$ according to Algorithm~$\mathsf{Prc}$.
    By the induction hypothesis, the existence of the SWD path for $B(a)$ can be determined.
    Thus, the existence of the SWD path of $A(a)$ can also be determined by $S_{\textit{\!\tiny rch}}$.

\item[Case~2] $R(a,b),B(b)\rightarrow A(a)$ where $\exists R.B\sqsubseteq A\in\mathcal{T}$.
    According to Algorithm~$\mathsf{Prc}$ and the fact that $R(a,b)$ has to be derived,
    we have that \texttt{rch}$(B(b),$ $A(a))$ is in $S_{\textit{\!\tiny rch}}$.
    By the induction hypothesis, the existence of the SWD path for $B(b)$ can be determined.
    Thus, the existence of the SWD path of $A(a)$ can also be determined.
\end{enumerate}

@@

\begin{enumerate}[leftmargin=12ex]
\item[Case~3] $A_1(a),A_2(a)\rightarrow A(a)$ where $A_1\sqcap A_2\sqsubseteq A\in\mathcal{T}$.
    If $A_1(a)$ is derived where $A_1$ is a simple concept,
    we have that \texttt{rch}$(A_2(a),A(a))$ is in $S_{\textit{\!\tiny rch}}$
    according to Algorithm~$\mathsf{Prc}$.
    By the induction hypothesis, the existence of the SWD path for $A_2(a)$ can be determined.
    Thus, the existence of the SWD path of $A(a)$ can also be determined.
    It is similar for the case where $A_2(a)$ is derived and $A_2$ is a simple concept.
\end{enumerate}

@@

\begin{enumerate}[leftmargin=12ex]
\item[Case~4] $S(a,b)\rightarrow R(a,b)$ where $\mathcal{O}\models S\sqsubseteq_* R$.
    We have that \texttt{rch}$(S(a,b),R(a,b))$ is in $S_{\textit{\!\tiny rch}}$
    according to Algorithm~$\mathsf{Prc}$.
    By the induction hypothesis, the existence of the SWD path for $S(a,b)$ can be determined.
    Thus, the existence of the SWD path of $R(a,b)$ can also be determined.

\item[Case~5] $S(b,a)\rightarrow R(a,b)$ where $\mathcal{O}\models S\sqsubseteq_* R^-$.
    We have that \texttt{rch}$(S(b,a),R(a,b))$ is in $S_{\textit{\!\tiny rch}}$
    according to Algorithm~$\mathsf{Prc}$.
    By the induction hypothesis, the existence of the SWD path for $S(b,a)$ can be determined.
    Thus, the existence of the SWD path of $R(a,b)$ can also be determined.

\item[Case~6] $R_1(a,c),R_2(c,b)\rightarrow R(a,b)$ where $R_1\circ R_2\sqsubseteq R\mathcal{R}$.
    If $R_1(a,c)$ is derived where $R_1$ is a simple role,
    we have that \texttt{rch}$(R_2(c,b),R(a,b))$ is in $S_{\textit{\!\tiny rch}}$
    according to Algorithm~$\mathsf{Prc}$.
    By the induction hypothesis, the existence of the SWD path for $R_2(c,b)$ can be determined.
    Thus, the existence of the SWD path of $R(a,b)$ can also be determined.
    It is similar for the case where $R_2(c,b)$ is derived and $R_2$ is a simple role.
\end{enumerate}
